---
title: "New developments in r-spatial"
date: "May 11, 2017"
output: html_document
---
@[edzerpebesma](https://twitter.com/edzerpebesma)

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

View these [slides](https://htmlpreview.github.io/?https://github.com/edzer/gsif/blob/master/gsif.html) or their [source](https://github.com/edzer/gsif)

## New developments in or around R

* tidyverse
* github
* blogs, twitter, [reprex](https://github.com/jennybc/reprex)
* data science

## r-spatial

* http://r-spatial.org/ and https://github.com/r-spatial/ not to be confused with
* http://rspatial.org/
* http://github.com/r-spatial
* sf succeeding sp, sf-verse succeeding [sp-verse](https://twitter.com/csgillespie/status/854438264538845184).
* calls for follow up on raster: [velox](https://github.com/hunzikp/velox), [fasterize](https://github.com/ecohealthalliance/fasterize), [tidy raster](https://github.com/ropensci/unconf17/issues/54), [stars](https://github.com/edzer/stars/blob/master/PROPOSAL.md)

## sf

Simple features are a standard for feature (mainly points, lines, polygons) data. The standard

* includes 17 geometry types,
* dominant are: `POINT`, `MULTIPOINT`, `LINESTRING`, `MULTILINESTRING`, `POLYGON`, `MULTIPOLYGON`, `GEOMETRYCOLLECTION`
* defines points in 2D, 3D, possibly with an `M` (measured) coordinate
* assumes simple (linear) interpolation between points
* has polygons as a single outer ring followed by zero or more inner rings (holes)
* has a text encoding (WKT), such as `LINESTRING(0 0,0 1,1 1)`
* has a binary encoding (WKB), used for I/O to libraries and storage in database blobs

Who use simple features? Everyone:

* GDAL
* GEOS
* lwgeom
* GeoJSON is a constrained set (epsg 4326, 7 types, polygons not crossing +/- 180E/W)
* spatial databases: PostGIS, SQLite, spatialite, GeoPackage, MySQL, ...
* ESRI's ArcGIS
* GeoSPARQL

## was anything wrong with `sp`?

No, but

* it does not implement simple features, but rather that what was around in 2003 (shapefiles)
* roundtripping simple features potentially leads to data loss
* its definition of polygon holes and the outer rings to which they belong is ambiguous and convoluted
* it doesn't work easily with tidyverse, `Spatial` objects are not `data.frame`s

## Package `sf`:

```{r}
library(sf)
```
implements simple features in three levels: as single geometries:
```{r}
(p1 = st_point(c(0,1)))
(p2 = st_point(c(2,3)))
(l = st_linestring(rbind(c(0,0), c(0,1), c(1,1))))
str(p1)
str(l)
```
as set of simple features _geometries_:
```{r}
(p = st_sfc(p1, p2, crs = 4326))
```
as simple features: objects with geometry
```{r}
(st = st_sf(soil_type = c("podzol", "fluvisol"), elevation = c(55, 12), geom = p))
class(st)
```
Note that `sf` objects _are_ `data.frame`s.

This model essentially mirrors that of a geometry column of spatial databases like PostGIS. We can deal with more than one geometry column.

## How does this work? Enter *list-columns*

```{r}
df = data.frame(soil_type = c("podzol", "fluvisol"), elevation = c(55, 12))
df$listcolumn = list(1:2, 1:5)
df
```
Now add a geometry list column:
```{r}
df$geom = p
df
class(df)
```
we still have a `data.frame`; so we do
```{r}
s = st_as_sf(df)
class(s)
s
```

# Now polygons
```{r}
nc = st_read(system.file("gpkg/nc.gpkg", package="sf"))
plot(nc[1])
```
but also
```{r}
library(ggplot2)
nc$geometry = nc$geom
ggplot(nc) + geom_sf(aes(fill = AREA))
```
```{r}    
# setting the aspect ratio correctly.
library(maps)
world1 <- sf::st_as_sf(map('world', plot = FALSE, fill = TRUE))
ggplot() + geom_sf(data = world1)
world2 <- sf::st_transform(
    world1,
    "+proj=laea +y_0=0 +lon_0=155 +lat_0=-90 +ellps=WGS84 +no_defs"
)
ggplot() + geom_sf(data = world2)
```
See also DE-9IM: https://en.wikipedia.org/wiki/DE-9IM
```{r}
st_rook = function(a, b = a) st_relate(a, b, pattern = "F***1****")
library(dplyr)
nc <- nc %>% mutate(NB_ROOK = st_rook(.))
nc[1:3,]
```